Learning and comparing functional connectomes across subjects

### 2013-01-01 paper.tex
iter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Neuroimage-like layout
\documentclass[5p]{elsarticle}
% For kindle
%\documentclass[1p,12pt]{elsarticle}
%\usepackage{geometry}
%\geometry{a6paper,hmargin={.2cm,.2cm},vmargin={1cm,1cm}}
% End kindle
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{bm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{url}
\usepackage[breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
\usepackage[table]{xcolor}

@@

% General introduction
Functional connectivity captures the synchronization of distant
neural systems via correlations in brain activity
\cite{biswal1995,cordes2000}. Given that high-level
function emerges from the interaction of specialized units, functional
connectivity is an essential part of the description of brain function,
complementary to the localizationist picture emerging from the systematic
mapping of regions recruited in tasks. However, while there exists a
well-defined standard analysis framework for activation mapping that enables
statistically-controlled comparisons across subjects \cite{friston1995},
group-level analysis of functional connectivity still faces many open
methodological challenges. Giving a picture of a single subject's 
functional connectivity is by itself not straightforward, as the brain
comprises a myriad of interacting subsystems and its connectivity must 
be decomposed in
simplified and synthetic representations. An important view of brain
connectivity is that of distributed functional networks depicted by their
spatial maps \cite{fox2005}. Another no less important and complementary
view is that of connections linking localized functional modules depicted
as a graph \cite{bullmore2009}. This representation of brain connectivity
is often called the functional connectome \cite{sporns2005} and is the
focus of intense worldwide research efforts as it holds promises of new
insights in cognition and pathologies
\cite{greicius2008b,biswal2010,fox2010}.

% Give the scope of the paper and narrow it
The purpose of this paper is to review methodological progress in the
estimation and comparison of functional connectomes. It does not attempt
to be exhaustive, as the field is wide and moving rapidly, but details
specific tools and guidelines that, in the experience of the authors,
lead to controlled and powerful inter-subject comparisons. The paper is
focused on functional connectomes, in contrast to structural connectomes,
as the inference of functional connectivity from fluctuations in the
signal implies important specificities in the statistical modeling. While 
the notion of functional
connectome is often associated with the study of resting-state
\cite{biswal2010}, the methods presented in this paper are also relevant
for task-based studies. On the other hand, the paper has a focus on fMRI;
although the core concepts presented can be applied to MEG or EEG
\cite{stam2004}, additional specific problems such as source
reconstruction must be considered \cite{schoffelen2009}.

% Define terms
By ``functional connectome'', here we specifically denote a graph
representing functional interactions in the brain, where the term
``graph'' is taken in its mathematical sens: a set of \emph{nodes}
connected together by \emph{edges}. A graph can be weighted or not, and
is completely equivalent to its \emph{adjacency matrix}, a symmetric
matrix tabulating the connection weights between each pair of nodes.
Functional-connectivity graphs can be used to represent evoked activity,
as in task-response studies \cite{mcintosh2000}, as well as
ongoing activity, present in the absence of specific tasks or in 
the background during task and often studied in so-called \emph{resting-state}
experiments \cite{raichle2010}. Another important notion that arises from
the study of distributed modes of brain function is that of specialized
functional networks\footnote{In neuroimaging, the term network is
sometimes used to denote a graph of brain function. To disambiguate the
notion of segregated spatial mode \cite{fox2005} from that of
connectivity graphs, we will purposely restrict its usage in this
paper.} \cite{fox2005}. With our definition of the functional connectome, functional
networks are not directly building blocks of the connectome but appear
as a consequence of the graphical structure
\cite{varoquaux2010c,varoquaux2012}.

@@

The choice of regions of interests (ROIs) that defines the nodes of the
graphs can be very important both in the estimation of connectomes and
for group comparison \cite{wang2009}. Unsurprisingly, simulations have
shown that extracting signal from ROIs that did not match functional
units would lead to incorrect graphs estimated \cite{smith2011}.
%
Different strategies to define suitable ROIs coexist. While dense parcellations approaches cover
a large fraction of the brain \cite{achard2006, varoquaux2010c,
wang2009}, this coverage can be traded off to focus on some specific
regions, in favor of increased functional specificity and thus better
differentiation across networks \cite{greicius2003, fair2009,
varoquaux2010b}. In addition, while
ROIs are most often defined as a hard selection of voxels, it is also
possible to use a \emph{soft} definition, attributing weights as with
probabilistic atlases, or spatial maps of functional networks extracted
from techniques such as independent component analysis (ICA).

@@

\paragraph{FMRI-based function definition}
%
Defining regions directly from the fMRI signal brings many benefits.
First, it can capture subject-specific functional information. Second, it
adapts to the signal at hand and its limitations, such as
EPI-specific warps or vascular and movement artifacts that are isolated
in ICA-like approaches. The simplest approach to define task-specific
regions is to use activation maps derived from standard GLM-based
analysis in a task-driven study (see for instance \cite{poldrack2011}).
Regions are extracted by thresholding the maps, or using balls around the
activation peaks. For resting-state studies, unsupervised multivariate
analysis techniques are necessary. Clustering approaches extract
full-brain parcellations \cite{craddock2011, bellec2010, yeo2011,
thirion2006}, and have been shown to extract well-known functional
structures from rest data. Alternatively, decompositions
methods, such as ICA \cite{beckmann2004}, can unmix linear
combinations of multiple effects and separate out overlapping spatial maps
capturing functional networks or confounding effects. At high model
order, ICA maps define a functional parcellation \cite{kiviniemi2009}.
Extracting regions from these maps requires additional effort as they can
display fragmented spatial features and structured background noise, but
incorporating sparsity and spatial constraints in the decomposition
techniques leads to contrasted maps that outline many different
structures \cite{varoquaux2012} (see fig.\,\ref{fig:parcellations}).

@@

The concept of functional connectivity has be called elusive
\cite{horwitz2003}: it has many mathematical instantiations although in
essence they all strive to extract simple statistics from functional
imaging to characterize synchrony and communication between large
ensembles of neurons. Here we choose to focus on second order statistics
that can be related to Gaussian models, the simplest of which being the
correlation matrix of the signals of the different ROIs.

@@

\paragraph{Signal extraction}
%
Given ROIs, the next step is to extract from them a summary measure. To
study \emph{on-going} activity, \emph{e.g.}\ with rest data, signal
extraction is performed from the EPI time series; however, to study
\emph{evoked} activity with task-driven studies it can be beneficial to
run a first-level analysis, enforcing specificity of the measure
extracted to the task. With slow event-related designs, task-specific
functional connectivity can be captured in trial-to-trial fluctuations in
the BOLD response, estimated using a GLM analysis with one regressor per
trial \cite{grillon2012,rissman2004}. This approach, known as beta-series
regression, has been adapted for rapid event-related designs, using
multiple GLMs to optimize deconvolution of each trial \cite{mumford2012}.
With rest datasets, it is important to regress out time series capturing
movement and other sources of structured noise, to separate on-going
activity from confounding signals. In addition to removing movement
parameters estimated during preprocessing, removing linear trends, signal
extracted from the white matter and the CSF \cite{chang2009}, as well as
using the CompCor \cite{behzadi2007} procedure gives more contrasted
correlation matrices that outline functional structures better
(fig.\,\ref{fig:correlation_matrices}). Filtering high frequencies is
often recommended, based on the initial observation that neural-related
signals are observed below 0.1\,Hz \cite{cordes2001,biswal1995}. While
high-pass and low-pass filtering decrease the impact of some confounds,
recent studies have shown that information on neural processes is present
in the full spectrum of frequencies observed
\cite{smith2012,vanoort2012}. Regressing out a good choice of confound
signals is more specific than frequency filter, and in our experience
gives more contrasted correlation matrices\footnote{Note that naive use
of filtering can induce spurious correlations \cite{davey2012}.}.

@@

\caption{
Different inverse-covariance matrices estimates corresponding to
Fig.\,\ref{fig:correlation_matrices} -- \textbf{Left}: group-sparse estimate
using the $\ell_{21}$ estimator \cite{varoquaux2010c}.
The insert shows the
connections restricted to a few major nodes. -- \textbf{Upper
right}: non-sparse estimate: inverse of the sample correlation matrix. --
\textbf{Lower right}: sparse estimate using the Graph Lasso
\cite{friedman2008}.
\label{fig:icov_estimators}
}
\end{figure}
%---------------------------------------------------------------------------}}}

@@

For the problem of recovering the functional-connectivity
\emph{structure}, \emph{i.e.}\ finding which region is connected to
which, sparse inverse covariance estimators have been found to be
efficient \cite{varoquaux2010c, smith2011}. The intuition for relying on
inverse covariance rather than correlation stems from that fact that
standard correlation (marginal correlation) between two variables $a$ and
$b$ also capture the effects of other variables: strong correlation of
$a$ and $b$ with a third variable $c$ will induce a correlation between
$a$ and $b$. On the opposite, the inverse covariance\footnote{Covariance
and correlation matrices differ simply by the fact that a covariance
matrix captures the amplitude of a signal, via its variance, while a
correlation matrix is computed on normalized signals.}
matrix captures
partial correlations, removing the effect of other variables
\cite{marrelec2006a}. In the small sample limit, this removal is
challenging from the statistical standpoint, and assumption of sparsity,
\emph{i.e.}\ that only few variables need to be considered at a time, is
important to estimate a good inverse covariance. Various estimation
strategies exist for sparse inverse covariance, and have an impact on the
resulting networks \cite{varoquaux2012,varoquaux2010c}. The
Graph Lasso ($\ell_1$-penalized maximum-likelihood estimator)
\cite{friedman2008} is in general a good approach for structure recovery. In group studies,
the $\ell_{21}$ estimator \cite{varoquaux2010c,honorio2012} is useful to
impose a common sparsity structure across different subjects and achieve
better recovery of this common structure. Simply put, these approaches
are necessary because estimation noise creates a background structure
(see Fig.\,\ref{fig:icov_estimators}); however, unlike in a univariate situation, the
parameters are not independent, and the spurious background connections
degrade the estimation of the actual connections. The sparse estimators
make a compromise between imposing simpler models, \emph{i.e.}\ with less
connections, and fitting well the data. This compromise is set via a
regularization parameter which controls the sparsity of the estimate. A
good rule of thumb to choose this parameter is via cross-validation
\cite{varoquaux2010c}.

@@

\paragraph{Network structure extracted}
%
The correlation matrices and inverse-covariance matrices that we extract
contain a lot of information on the functional structure of the brain.
First, the correlation matrix (Fig.\,\ref{fig:correlation_matrices})
shows blocks of coactivating regions that can be interpreted as
large-scale functional networks, such as the default mode network. Note
that the split in networks is not straightforward. Different ordering of the
nodes will reveal different networks. Indeed, because of the presence of
hubs and interleaved networks, the picture in terms of segregated networks
is not sufficient to explain full-brain connectivity
\cite{varoquaux2012}. The connectivity matrices, correlation matrix or
inverse-covariance matrix, can be represented as graphs: nodes connected
by weighted edges (inserts on Fig.\,\ref{fig:correlation_matrices} and
Fig.\,\ref{fig:icov_estimators}). The inverse-covariance matrix,
capturing partial correlations, appears then as extracting a
\emph{backbone} or \emph{core} of the graph. While such structure have
been used as a way to summarize anatomical brain connectivity graphs
\cite{hagmann2008}, here they have a clear-cut meaning with regards to
the BOLD signal: they give the conditional independence structure between
the regions \cite{varoquaux2012}. In other words, regions $a$ and $b$ are
not connected if the signal that they have in common can be explained by
a third region $c$. In this light, the choice of nuisance regressors to
remove confounding common signal is less critical with partial
correlations than with correlations. Indeed, while with
correlation matrices regressing out the global mean has a drastic
(Fig.\,\ref{fig:correlation_matrices} upper right and lower right), on
inverse covariance it only changes very slightly the resulting matrices
(Fig.\,\ref{fig:icov_regressors}). There have been debates on whether to
regress out certain signals, such as the global mean,  as it induces
negative correlations \cite{murphy2009,chang2009,fox2009}, and these may seem
surprising: one network is doing the opposite of the other. However,
correlation between two signals only takes its meaning with the
definition of a baseline. A simple picture to explain anti-correlations
between two regions is the presence of a third region, mediating the
interactions. Using this third region as a baseline would amount to
estimate partial correlations in the whole system. Using 
inverse-covariance matrices or partial correlations to understand brain
connectivity makes the interpretation in terms of interactions between
brain regions easier and more robust to the choice of confounds.

@@

In this section, we consider the problem of comparing functional
connectivity across subjects or across conditions.

@@

\paragraph{Mass-univariate approaches}
%
The most natural approach is to apply a linear model to each coefficient of
the connectivity matrices \cite{lewis2009,grillon2012}. This
approach is similar to the second-level analysis used in mass-univariate
brain mapping, and gives rise to many of the well-known techniques
used in such a context, such as the definition of a second-level design,
with possibly the inclusion of confounding effects, and statistical tests
(T tests or F tests) on contrast vector. Importantly, in order to work
with Gaussian-distributed variables, it is necessary to apply a Fisher
Z transform\footnote{See
\url{http://en.wikipedia.org/wiki/Fisher_transformation} or
\cite{anderson1958}, section 4.2.3 for mathematical arguments} to the
correlations. Note that
in these settings, the Ledoit-Wolf estimator \cite{ledoit2004} is
often a good choice to estimate the correlation matrix, as it is
parameter-free and gives good estimation performance without imposing any
restrictions on the data.
%
For hypothesis testing,
correcting for multiple comparison can severely limit statistical
power, as the number of tests performed scales as the square of the
number of regions used. Controlling for the false discovery rate (FDR)
mitigates this problem. Alternatively, as the assumption underlying the
Benjamini-Hochberg procedure \cite{benjamini1995} for the FDR can easily
be broken, non-parametric permutations-based approaches give reliable
approaches. In particular, the max-T procedure \cite{ge2003,nichols2001}
is interesting to avoid the drastic Bonferroni correction when
controlling for multiple comparison in family-wise error rate.

@@

\caption{Inter subject variability. Note that this is variability
occurring in a healthy population at rest, in other words non specific
variability -- \textbf{a}: single-subject
correlation matrices for different subjects -- \textbf{b}:
Corresponding Z-score (effect / standard deviation) of the difference
between a subject and the remaining others -- \textbf{c}:
single-subject precision matrices -- \textbf{d}: Corresponding Z-score
for the precision matrices -- \textbf{e}:
Corresponding Z-score for the subject residuals, as defined in 
\cite{varoquaux2010b}.
\label{fig:inter_subject}}
\end{figure}
%---------------------------------------------------------------------------}}}

\paragraph{Accounting for distributed variability}
%
A specific challenge of connectivity analysis is that the connectivity
strength between different regions tend to covary. For instance, with
resting-state data, functional networks comprising many nodes can appear
as more or less connected across subjects (see for instance
Fig.\,\ref{fig:inter_subject}.a, showing variability in a control
population at rest). In other words, non-specific variability is
distributed accross the connectivity graph, and it is structured by the
graph itself. This observation brings the natural question of whether
second-level analysis should be performed on correlation matrices,
inverse-covariance matrices, or another parametrization that would
disentangle effects and give unstructured (white) residuals. While
inverse-covariance matrices show less distributed fluctuations than
correlation matrices, they capture a lot of background noise, as partial
correlations are intrinsically harder to estimate. Preliminary work
\cite{varoquaux2010b} suggests to perform statistics on residuals of
a parametrization intermediate between correlation matrix and inverse
covariance matrix, that decouples effects and noise.

@@

\paragraph{Network integration}
%
Marrelec \emph{et al.}\
\cite{marrelec2008} introduce the use of entropy and mutual information
as a measure of network integration\footnote{See \cite{varoquaux2010c}
for simplified formulas for network integration and mutual information.}.
Gaussian entropy can be seen as a simple metric to generalize correlation
or variance to multiple nodes (see \cite{anderson1958} \S2.5.2 and
\S7.5). Indeed, let us consider 3 nodes: $a$, $b$ and $c$. Their
correlation structure is captured by three correlation coefficients:
$\rho_{ab}$, $\rho_{bc}$ and $\rho_{ac}$. Summarizing these by their
mean, as it might seem natural, discards the relationship between the
signals, while using the integration tells us how much two signals can be
combined to form the third (see Fig.\,\ref{fig:integration}). 
Cross-entropy --or mutual information--
\cite{marrelec2008} measures the amount of cross-talk between two
systems in a similar way as Gaussian entropy to used to measure the
integration of a brain
system. The functional-connectivity structure, or its representation in
form of a correlation matrix, can thus be characterized via the
integration and cross-talk of some of its subs-systems. This approach
gives a simplified representation with a small number of metrics that can
be compared across subjects.

@@

Predictive modeling is concerned with learning (or fitting) a model to
draw conclusions on new and unseen data \cite{pereira2009}. Typically on
brain imaging data, it opens the door to learning diagnosis or prognosis
markers \cite{stonnington2010}. However, beyond the practical use of
prediction, these models can also be useful to uncover mechanism of brain
function \cite{norman2006}. Both the fact that information \emph{can} be
drawn to generalize predictions to new data and \emph{how} or
\emph{where} it lies informs us on brain function. Multivariate
predictive models
are attractive in connectomics because they are sensitive to dependencies
between features and avoid the need to correct for multiple comparisons
since the significance of an entire pattern is evaluated using a single
statistical test. From the perspective of medical applications, using
resting-state connectivity as a biomarker is particularly appealing as it
can be applied to impaired patients \cite{greicius2008b}. Predictive
modeling techniques have been successfully applied to identify
connectome-based biomarkers of Alzheimer's disease
\cite{stonnington2010}, depression \cite{craddock2009}, schizophrenia
\cite{cecchi2009}, autism \cite{anderson2011}, ADHD \cite{zhu2008}, aging
\cite{dosenbach2010}, as well as to classify mental operations
\cite{richiardi2011,shirer2012}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Neuroimage-like layout
%\documentclass[5p]{elsarticle}
% For kindle
\documentclass[1p,12pt]{elsarticle}
\usepackage{geometry}
\geometry{a6paper,hmargin={.2cm,.2cm},vmargin={1cm,1cm}}
% End kindle
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{bm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{url}
\usepackage[breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
\usepackage[table]{xcolor}

@@

% General introduction
Functional connectivity captures the synchronization of distant
neural systems via correlations in brain activity
\cite{biswal1995,cordes2000}. Given that high-level
function emerges from the interaction of specialized units, functional
connectivity is an essential part of the description of brain function,
that complements the localizationist picture emerging from the systematic
mapping of regions recruited in tasks. However, while there exists a
well-defined standard analysis framework for activation mapping that enables
statistically-controlled comparisons across subjects \cite{friston1995},
group-level analysis of functional connectivity still faces many open
methodological challenges. Giving a picture of a single subject's 
functional connectivity is by itself not straightforward, as the brain
comprises a myriad of interacting subsystems and its connectivity must 
be decomposed in
simplified and synthetic representations. An important view of brain
connectivity is that of distributed functional networks depicted by their
spatial maps \cite{fox2005}. Another no less important and complementary
view is that of connections linking localized functional modules depicted
as a graph \cite{bullmore2009}. This representation of brain connectivity
is often called the functional connectome \cite{sporns2005} and is the
focus of intense worldwide research efforts as it holds promises of new
insights in cognition and pathologies
\cite{greicius2008b,biswal2010,fox2010}.

% Give the scope of the paper and narrow it
The purpose of this paper is to review methodological progress in the
estimation and comparison of functional connectomes. It does not attempt
to be exhaustive, as the field is wide and moving rapidly, but details
specific tools and guidelines that, in the experience of the authors,
lead to controlled and powerful inter-subject comparisons. The paper is
focused on functional connectomes in contrast to structural connectomes,
as the inference of functional connectivity from fluctuations in the
signal implies important specificities in the statistical modeling. While 
the notion of functional
connectome is often associated with the study of resting-state
\cite{biswal2010}, the methods presented in this paper are also relevant
for task-based studies. On the other hand, the paper has a focus on fMRI;
although the core concepts presented can be applied to MEG or EEG
\cite{stam2004}, additional specific problems such as source
reconstruction must be considered \cite{schoffelen2009}.

% Define terms
By ``functional connectome'', here we specifically denote a graph
representing functional interactions in the brain, where the term
``graph'' is taken in its mathematical sens: a set of \emph{nodes}
connected together by \emph{edges}. A graph can be weighted or not, and
is completely equivalent to its \emph{adjacency matrix}, a symmetric
matrix tabulating the connection weights between each pair of nodes.
Functional-connectivity graphs are used to represent evoked activity,
as in task-response studies \cite{mcintosh2000}, as well as
ongoing activity, present in the absence of specific tasks or in 
the background during task and often studied in so-called \emph{resting-state}
experiments \cite{raichle2010}. Another important notion that arises from
the study of distributed modes of brain function is that of specialized
functional networks\footnote{In neuroimaging, the term network is
sometimes used to denote a graph of brain function. To disambiguate the
notion of segregated spatial mode \cite{fox2005} from that of
connectivity graphs, we will purposely restrict its usage in this
paper.} \cite{fox2005}. With our definition of the functional connectome, functional
networks are not directly building blocks of the connectome but appear
as a consequence of the graphical structure
\cite{varoquaux2010c,varoquaux2012}.

@@

The choice of regions of interests (ROIs) that define the nodes of the
graphs can be very important both in the estimation of connectomes and
for group comparison \cite{wang2009}. Unsurprisingly, simulations have
shown that extracting signal from ROIs that did not match functional
units would lead to incorrect graphs estimated \cite{smith2011}.
%
Different strategies to define suitable ROIs coexist. While dense parcellations approaches cover
a large fraction of the brain \cite{achard2006, varoquaux2010c,
wang2009}, this coverage can be traded off to focus on some specific
regions, in favor of increased functional specificity and thus better
differentiation across networks \cite{greicius2003, fair2009,
varoquaux2010b}. In addition, while
ROIs are most often defined as a hard selection of voxels, it is also
possible to use a \emph{soft} definition, attributing weights as with
probabilistic atlases, or spatial maps of functional networks extracted
from techniques such as independent component analysis (ICA).

@@

\paragraph{FMRI-based function definition}
%
Defining regions directly from the fMRI signal brings many benefits.
First, it can capture subject-specific functional information. Second, it
adapts to the signal at hand and its limitations, such as
EPI-specific warps or vascular and movement artifacts that are isolated
in ICA-like approaches. The simplest approach to define task-specific
regions is to use activation maps derived from standard GLM-based
analysis in a task-driven study (see for instance \cite{poldrack2011}).
Regions are extracted by thresholding the maps, or using balls around the
activation peaks. For resting-state studies, unsupervised multivariate
analysis techniques are necessary. Clustering approaches extract
full-brain parcellations \cite{craddock2011, bellec2010, yeo2011,
thirion2006}, and have been shown to segment well-known functional
structures from rest data. Alternatively, decompositions
methods, such as ICA \cite{beckmann2004}, can unmix linear
combinations of multiple effects and separate out overlapping spatial maps
capturing functional networks or confounding effects. At high model
order, ICA maps define a functional parcellation \cite{kiviniemi2009}.
Extracting regions from these maps requires additional effort as they can
display fragmented spatial features and structured background noise, but
incorporating sparsity and spatial constraints in the decomposition
techniques leads to contrasted maps that outline many different
structures \cite{varoquaux2012} (see fig.\,\ref{fig:parcellations}).

@@

The concept of functional connectivity has be called elusive
\cite{horwitz2003}: it has many mathematical instantiations although in
essence they all strive to extract simple statistics from functional
imaging in order to characterize synchrony and communication between large
ensembles of neurons. Here we choose to focus on second order statistics
that can be related to Gaussian models, the simplest of which being the
correlation matrix of the signals of the different ROIs.

@@

\paragraph{Signal extraction}
%
Given ROIs, the next step is to extract from them a summary measure. To
study \emph{on-going} activity, \emph{e.g.}\ with rest data, signal
extraction is performed from the EPI time series; however, to study
\emph{evoked} activity with task-driven studies it can be beneficial to
run a GLM-based first-level analysis, enforcing specificity of the measure
extracted to the task. With slow event-related designs, task-specific
functional connectivity can be captured in trial-to-trial fluctuations in
the BOLD response, estimated using a GLM analysis with one regressor per
trial \cite{grillon2012,rissman2004}. This approach, known as beta-series
regression, has been adapted for rapid event-related designs, using
multiple GLMs to optimize deconvolution of each trial \cite{mumford2012}.
With rest datasets, it is important to regress out time series capturing
movement and other sources of structured noise, to separate on-going
activity from confounding signals. In addition to removing movement
parameters estimated during preprocessing, removing linear trends, signal
extracted from the white matter and the CSF \cite{chang2009}, as well as
using the CompCor \cite{behzadi2007} procedure gives more contrasted
correlation matrices that outline functional structures better
(fig.\,\ref{fig:correlation_matrices}). Filtering high frequencies is
often recommended, based on the initial observation that neural-related
signals are observed below 0.1\,Hz \cite{cordes2001,biswal1995}. While
high-pass and low-pass filtering decrease the impact of some confounds,
recent studies have shown that information on neural processes is present
in the full spectrum of frequencies observed
\cite{smith2012,vanoort2012}. Regressing out a good choice of confound
signals is more specific than frequency filter, and in our experience
gives more contrasted correlation matrices\footnote{Note that naive use
of filtering can induce spurious correlations \cite{davey2012}.}.

@@

\caption{
Different inverse-covariance matrices estimates corresponding to
fig.\,\ref{fig:correlation_matrices} -- \textbf{Left}: group-sparse estimate
using the $\ell_{21}$ estimator \cite{varoquaux2010c}.
The insert shows the
connections restricted to a few major nodes. -- \textbf{Upper
right}: non-sparse estimate: inverse of the sample correlation matrix. --
\textbf{Lower right}: sparse estimate using the Graph Lasso
\cite{friedman2008}.
\label{fig:icov_estimators}
}
\end{figure}
%---------------------------------------------------------------------------}}}

@@

For the problem of recovering the functional-connectivity
\emph{structure}, \emph{i.e.}\ finding which region is connected to
which, sparse inverse covariance estimators have been found to be
efficient \cite{varoquaux2010c, smith2011}. The intuition for relying on
inverse covariance rather than correlation stems from that fact that
standard correlation (marginal correlation) between two variables $a$ and
$b$ also capture the effects of other variables: strong correlation of
$a$ and $b$ with a third variable $c$ will induce a correlation between
$a$ and $b$. On the opposite, the inverse covariance\footnote{Covariance
and correlation matrices differ simply by the fact that a covariance
matrix captures the amplitude of a signal, via its variance, while a
correlation matrix is computed on normalized signals.}
matrix captures
partial correlations, removing the effect of other variables
\cite{marrelec2006a}. In the small sample limit, this removal is
challenging from the statistical standpoint, and an assumption of sparsity,
\emph{i.e.}\ that only few variables need to be considered at a time, is
important to estimate a good inverse covariance. Various estimation
strategies exist for sparse inverse covariance, and have an impact on the
resulting networks \cite{varoquaux2012,varoquaux2010c}. The
Graph Lasso ($\ell_1$-penalized maximum-likelihood estimator)
\cite{friedman2008} is in general a good approach for structure recovery. In group studies,
the $\ell_{21}$ estimator \cite{varoquaux2010c,honorio2012} is useful to
impose a common sparsity structure across different subjects and achieve
better recovery of this common structure. Simply put, these approaches
are necessary because estimation noise creates a background structure
(see fig.\,\ref{fig:icov_estimators}); however, unlike in a univariate situation, the
parameters are not independent, and the spurious background connections
degrade the estimation of the actual connections. The sparse estimators
make a compromise between imposing simpler models, \emph{i.e.}\ with less
connections, and fitting well the data. This compromise is set via a
regularization parameter which controls the sparsity of the estimate. A
good rule of thumb to choose this parameter is via cross-validation
\cite{varoquaux2010c}.

@@

\paragraph{Network structure extracted}
%
The correlation matrices and inverse-covariance matrices that we extract
contain a lot of information on the functional structure of the brain.
First, the correlation matrix (fig.\,\ref{fig:correlation_matrices})
shows blocks of coactivating regions that can be interpreted as
large-scale functional networks, such as the default mode network. Note
that the split in networks is not straightforward. Different ordering of the
nodes will reveal different networks. Indeed, because of the presence of
hubs and interleaved networks, the picture in terms of segregated networks
is not sufficient to explain full-brain connectivity
\cite{varoquaux2012}. The connectivity matrices, correlation matrix or
inverse-covariance matrix, can be represented as graphs: nodes connected
by weighted edges (inserts on fig.\,\ref{fig:correlation_matrices} and
fig.\,\ref{fig:icov_estimators}). The inverse-covariance matrix,
capturing partial correlations, appears then as extracting a
\emph{backbone} or \emph{core} of the graph. While such structure has
been used as a way to summarize anatomical brain connectivity graphs
\cite{hagmann2008}, here it has a clear-cut meaning with regards to
the BOLD signal: it gives the conditional independence structure between
the regions \cite{varoquaux2012}. In other words, regions $a$ and $b$ are
not connected if the signal that they have in common can be explained by
a third region $c$. In this light, the choice of nuisance regressors to
remove confounding common signal is less critical with partial
correlations than with correlations. Indeed, while with
correlation matrices regressing out the global mean has a drastic effect
(fig.\,\ref{fig:correlation_matrices} upper right and lower right), on
inverse covariance it only changes very slightly the resulting matrices
(fig.\,\ref{fig:icov_regressors}). There have been debates on whether to
regress out certain signals, such as the global mean,  as it induces
negative correlations \cite{murphy2009,chang2009,fox2009}, and these may seem
surprising: one network is doing the opposite of the other. However,
correlation between two signals only takes its meaning with the
definition of a baseline. A simple picture to explain anti-correlations
between two regions is the presence of a third region, mediating the
interactions. Using this third region as a baseline would amount to
estimate partial correlations in the whole system. Using 
inverse-covariance matrices or partial correlations to understand brain
connectivity makes the interpretation in terms of interactions between
brain regions easier and more robust to the choice of confounds.

@@

We now turn to the problem of comparing functional
connectivity across subjects or across conditions.

@@

\paragraph{Mass-univariate approaches}
%
The most natural approach is to apply a linear model to each coefficient of
the connectivity matrices \cite{lewis2009,grillon2012}. This
approach is similar to the second-level analysis used in mass-univariate
brain mapping, and gives rise to many of the well-known techniques
used in such a context, such as the definition of a second-level design,
with possibly the inclusion of confounding effects, and statistical tests
(T tests or F tests) on contrast vector. Importantly, in order to work
with Gaussian-distributed variables, it is necessary to apply a Fisher
Z transform\footnote{See
\url{http://en.wikipedia.org/wiki/Fisher_transformation} or
\cite{anderson1958} section 4.2.3 for mathematical arguments.} to the
correlations. Note that
in these settings, the Ledoit-Wolf estimator \cite{ledoit2004} is
often a good choice to estimate the correlation matrix, as it is
parameter-free and gives good estimation performance without imposing any
restrictions on the data.
%
For hypothesis testing,
correcting for multiple comparison can severely limit statistical
power, as the number of tests performed scales as the square of the
number of regions used. Controlling for the false discovery rate (FDR)
mitigates this problem. Alternatively, as the assumptions underlying the
Benjamini-Hochberg procedure \cite{benjamini1995} for the FDR can easily
be broken, non-parametric permutations-based tests give reliable
approaches. In particular, the max-T procedure \cite{ge2003,nichols2001}
is interesting to avoid the drastic Bonferroni correction when
controlling for multiple comparison in family-wise error rate.

@@

\caption{Inter subject variability. Note that this is variability
occurring in a healthy population at rest, in other words it is non specific
variability -- \textbf{a}: single-subject
correlation matrices for different subjects -- \textbf{b}:
Corresponding Z-score (effect / standard deviation) of the difference
between a subject and the remaining others -- \textbf{c}:
single-subject precision matrices -- \textbf{d}: Corresponding Z-score
for the precision matrices -- \textbf{e}:
Corresponding Z-score for the subject residuals, as defined in 
\cite{varoquaux2010b}.
\label{fig:inter_subject}}
\end{figure}
%---------------------------------------------------------------------------}}}

\paragraph{Accounting for distributed variability}
%
A specific challenge of connectivity analysis is that the connectivity
strength between different regions tend to covary. For instance, with
resting-state data, functional networks comprising many nodes can appear
as more or less connected across subjects (see for instance
fig.\,\ref{fig:inter_subject}.a, showing variability in a control
population at rest). In other words, non-specific variability is
distributed accross the connectivity graph, and it is structured by the
graph itself. This observation brings the natural question of whether
second-level analysis should be performed on correlation matrices,
inverse-covariance matrices, or another parametrization that would
disentangle effects and give unstructured (white) residuals. While
inverse-covariance matrices show less distributed fluctuations than
correlation matrices, they capture a lot of background noise, as partial
correlations are intrinsically harder to estimate. Preliminary work
\cite{varoquaux2010b} suggests to perform statistics on residuals of
a parametrization intermediate between correlation matrix and inverse
covariance matrix, that decouples effects and noise.

@@

\paragraph{Network integration}
%
Marrelec \emph{et al.}\
\cite{marrelec2008} introduce the use of entropy and mutual information
as a measure of network integration\footnote{See \cite{varoquaux2010c}
for simplified formulas for network integration and mutual information.}.
Gaussian entropy can be seen as a simple metric to generalize correlation
or variance to multiple nodes (see \cite{anderson1958} \S2.5.2 and
\S7.5). Indeed, let us consider 3 nodes: $a$, $b$ and $c$. Their
correlation structure is captured by three correlation coefficients:
$\rho_{ab}$, $\rho_{bc}$ and $\rho_{ac}$. Summarizing these by their
mean, as it might seem natural, discards the relationship between the
signals, while using the integration tells us how much two signals can be
combined to form the third (see fig.\,\ref{fig:integration}). 
Cross-entropy --or mutual information--
\cite{marrelec2008} measures the amount of cross-talk between two
systems in a similar way as Gaussian entropy is used to measure the
integration of a brain
system. The functional-connectivity structure, or its representation in
form of a correlation matrix, can thus be characterized via the
integration and cross-talk of some of its subs-systems. This approach
gives a simplified representation with a small number of metrics that can
be compared across subjects.

@@

Predictive modeling is concerned with learning (or fitting) a model to
draw conclusions on new and unseen data \cite{pereira2009}. Typically, on
brain imaging data, it opens the door to learning diagnosis or prognosis
markers \cite{stonnington2010}. However, beyond the practical use of
prediction, these models can also be useful to uncover mechanisms of brain
function \cite{norman2006}. Both the fact that information \emph{can} be
used to make predictions on new data, and \emph{how} it is used or
\emph{where} it lies informs us on brain function. Multivariate
predictive models
are attractive in connectomics because they are sensitive to dependencies
between features and avoid the curse of multiple comparisons
since the significance of an entire pattern is evaluated using a single
statistical test. From the perspective of medical applications, using
resting-state connectivity as a biomarker is particularly appealing as it
can be applied to impaired patients \cite{greicius2008b}. Predictive
modeling techniques have been successfully applied to identify
connectome-based biomarkers of Alzheimer's disease
\cite{stonnington2010}, depression \cite{craddock2009}, schizophrenia
\cite{cecchi2009}, autism \cite{anderson2011}, ADHD \cite{zhu2008}, aging
\cite{dosenbach2010}, as well as to classify mental operations
\cite{richiardi2011,shirer2012}.
